---
title: "Dog-SEP - preregistration"
author: "Nina & Radek"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default  
# csl: apa.csl
# bibliography: library.bib
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# packages
library(tidyverse)

# data
```

# Study Information

## Title

Dog breeds as indicators of neighbourhood socioeconomic position.

## Authorship

Schnyder, N.^1^ & Panczak, R.^2^ 

^1^ Schuepfheim, Switzerland

^2^ Wloki, Poland

## Description/Background (detailed)

Australians love their dogs. Despite the fact that buying and owning one can add up to large expenses. 

Tracking socioeconomic position (SEP) of people and neighbourhoods is not a trivial task and usually requires access to comprehensive and high quality data that are often sensitive in nature. In the absence of individual-level SEP measures (such as income, wealth and education) area based measures are often used. Such measures usually use one statistical variable or use more to construct an index and use it to describe SEP of the area.

Census variables have been widely used for this purpose with more recent attempts leaning towards more exotic measures derived from night-time lights, mobile phones or car registrations to name few.

## Research Questions and Hypotheses

We will attempt to answer the question if the dog breed ownership data can be used as proxy for SEP of suburbs using the city of Brisbane, Australia as an example.

We hypothesize that the cost of buying and maintaining the dog is a proxy for socioeconomic position of the owner. In the absence of individual level data we hypothesize that these associations will hold on the neighbourhood level. This will be indicated by higher amount of expensive dog breeds among all dogs in the neighbourhood. 

---

# Design Plan

## Study Type

Observational study.

## Study design

Cross-sectional study.

## Blinding

No blinding is involved in this study.

## Randomisation

Not applicable.

---

# Sampling Plan

## Existing data

### Dog ownership

`Dog ownership` dataset comes from the Brisbane City Council(BCC) [open data portal](https://www.data.brisbane.qld.gov.au/data/dataset/current-animal-related-permits/resource/c2823090-8a44-4c8b-9ed4-e988491b25f4). We will use `CARS - BIS - Open Data - Animal Permits Breed - 1 Jul 2019.csv` dataset that is closest in time to the `SEIFA` dataset. BBC's data portal does not provide earlier version of their data. Unfortunately, our attempts to contact BCC and obtain earlier version were not successful. In case the earlier dataset becomes available we will use that instead.

### Dog buying cost


### Dog insurance cost


### SEIFA

SEIFA indices are derived from 2016 census data and are available from [Austrlian Bureau of Statistics](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/2033.0.55.0012016?OpenDocument) (ABS).

We will search for associations of `Dog ownership` variables with all four indexes: 

  - The Index of Relative Socio-economic Disadvantage (IRSD); 
  - The Index of Relative Socio-economic Advantage and Disadvantage (IRSAD); 
  - The Index of Education and Occupation (IEO); 
  - The Index of Economic Resources (IER).
  
### Spatial data

We will use `State Suburbs ASGS Ed 2016 Digital Boundaries in ESRI Shapefile Format` from the [ABS](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.003July%202016?OpenDocument) as the primary definition of suburbs. Email correspondence from the BBC confirmed that these should match their data. Using this suburbs 

In case of inconsistencies, mismatches or name conflicts we will use `Locality boundaries - Queensland` data from the from Queensland Spatial Catalogue - [QSpatial](http://qldspatial.information.qld.gov.au/catalogue/custom/detail.page?fid={8F24D271-EE3B-491C-915C-E7DD617F95DC).

## Explanation of existing data and prior work based on this data set

We explored frequencies of variables in `Dog ownership` data in order to resolve what analyses could be feasible. We checked the level of details in coding of the `Animal: Breed` variable. We also checked the values of the `Application Location: Suburb` variable in order to determine what spatial units were used for reporting and match that to the `SEIFA` data.

## Data collection procedures and brief description of data

All data has been collected and existed prior to the design of the study.

## Sample size and data inclusion/exclusion

Not applicable.

## Sample size rationale

Not applicable.

## Stopping rule

Not applicable.

---

# Variables

## Manipulated variables

Not applicable.

## Measured variables 

### Dog ownership

`Animal: Breed` variable. 

`Application Location: Suburb` variable. 

### Dog cost


### Dog insurance cost


## Indices 

Four `SEIFA` indicators come in the form of indices. As recommended by ABS, we will use deciles of indices.

---

# Analysis Plan

## Transformations

### Dog ownership

In the first analysis, `Animal: Breed` variable will be recoded to binary using `Dog cost` dataset. 

`Application Location: Suburb` variable will then be used to aggregate the expensive and total amount of dogs in neighbourhood. 

These two variables will then be used to derive proportion.

In the second analysis `Animal: Breed` variable will be ... using `Dog insurance cost` dataset. 

## Missing Data

### Dog ownership

We will exclude cases with missing information on breed. 

### SEIFA

We will exclude neighbourhoods if one of the indices is missing.

## Statistical methods and models, including sampling weights

  

---

# References

# Template info 

This preregistration document is been based on the [OSF Prereg](https://osf.io/zab38/wiki/home/) R Markdown template by James Bartlett [available here](https://github.com/BartlettJE/BartlettJE.github.io/blob/master/RMarkdown-scripts/OSF%20preregistration%20template.Rmd) incorporating parts of the template for secondary data preregistration by [Weston and colleagues](https://osf.io/x4gzt/), see also [here](https://osf.io/jqxfz/).

