---
title: "Dog-SEP"
subtitle: "Dog cost"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

<!-- ------------------------------------------------------------ --> 

```{r r-setup, include = FALSE}
set.seed(12345)
options(scipen = 999)
options(max.print = "75")

library(pacman)
p_load(tidyverse, magrittr, scales,
       sjmisc, kableExtra,
       sf, tmap,
       correlation, see, modelbased)

tmap_mode("view") # makes map interactive

# library("tidylog")
# library("conflicted")
# for (f in getNamespaceExports("tidylog")) {
#   conflicted::conflict_prefer(f, "tidylog", quiet = TRUE)
# }
```

```{r knit-setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.width=8, fig.height=6, dpi=300, 
                      out.width="800px", out.height="600px")

knitr::opts_chunk$set(cache = FALSE,
                      prompt = FALSE,
                      tidy = FALSE,
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(width = 75)

mu <- Hmisc::markupSpecs$html
```

<!-- ------------------------------------------------------------ --> 

# Data 

Preparations in file `00.Rmd`

```{r}
dog_ownership_cost <- read_rds("data/permits/dog_ownership_cost.Rds") 

SSC <- read_rds("data/geo/SSC.Rds")
```

<!-- ------------------------------------------------------------ --> 

# All Brisbane dogs combined 

```{r, fig.width=8, fig.height=8}
my_colors <- RColorBrewer::brewer.pal(3, "BrBG")[c(1,3)]

dog_ownership_cost %>% 
  mutate(expensive = factor(expensive, levels = c(0, 1),
                            labels= c("No", "Yes"))) %>% 
  mutate(dog_breed = fct_lump(dog_breed, n = 50)) %>%
  group_by(dog_breed) %>% 
  mutate(total = n()) %>% 
  group_by(dog_breed, expensive) %>% 
  summarise(count = n(),
            total = first(total)) %>% 
  ggplot(aes(x = reorder(dog_breed, total) , 
             y = count,
             fill = expensive)) + 
  geom_col() +
  coord_flip() +
  ylab("Number of dogs") + xlab("") + 
  scale_fill_manual(values = my_colors) +
  theme_modern() +
  theme(panel.grid.major.x = element_line(colour = "grey80"),
        axis.text.y = element_text(size = 8))
```

## Dogs per capita

```{r}
dog_ownership_agg <- dog_ownership_cost %>% 
  group_by(SSC_NAME16) %>% 
  summarise(dogs_total = n(),
            dogs_exp = sum(expensive))

SSC %<>% 
  left_join(dog_ownership_agg)
```

Areas with no URP/SEIFA but having (small amount of) dogs:

```{r}
SSC %>% 
  st_drop_geometry() %>% 
  right_join(dog_ownership_agg) %>% 
  filter(is.na(SSC_CODE16)) %>% 
  select(SSC_NAME16, dogs_total)
```

Area with low URP and no dogs at all:

```{r}
SSC %>% 
  st_drop_geometry() %>% 
  left_join(dog_ownership_agg) %>% 
  filter(is.na(dogs_total)) %>% 
  select(SSC_NAME16, dogs_total, URP)
```

```{r}
SSC %<>% 
  left_join(dog_ownership_agg) %>%
  filter(!is.na(dogs_total)) %>%
  mutate(dogs_percap = (dogs_total / URP)*1000,
         dogs_exp_prop = dogs_exp / dogs_total)
```

Top 10 n'hoods:

```{r}
SSC %>% 
  st_drop_geometry() %>% 
  arrange(desc(dogs_percap)) %>% 
  select(SSC_NAME16, dogs_total, URP, dogs_percap) %>% 
  slice(1:10)
```

All on map: 

```{r}
tm_shape(SSC) +
  tm_bubbles(id = "SSC_NAME16", 
             size = "dogs_total", 
             col = "dogs_percap",
             palette = "BrBG",
             border.col = NA)
```

## Proportion of expensive dogs

```{r}
summary(SSC$dogs_exp_prop)
```

Top 10 n'hoods:

```{r}
SSC %>% 
  st_drop_geometry() %>% 
  arrange(desc(dogs_exp_prop)) %>% 
  select(SSC_NAME16, dogs_total, URP, dogs_exp_prop) %>% 
  slice(1:10)
```

```{r eval=FALSE, include=FALSE}
dog_ownership <- read_rds("data/permits/dog_ownership.Rds") 

dog_ownership %>% 
  filter(SSC_NAME16 == "Wacol") %>% 
  group_by(dog_breed) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  arrange(desc(n))

```

All on map: 

```{r}
tm_shape(SSC) +
  tm_bubbles(id = "SSC_NAME16", 
             size = "dogs_total", 
             col = "dogs_exp_prop",
             palette = "BrBG",
             border.col = NA)
```

<!-- ------------------------------------------------------------ --> 

# Association with SEIFA

## Functions

```{r}
seifa_means <- function (seifa_index) {
  
  myenc <- enquo(seifa_index)
  
  SSC %>% 
    st_drop_geometry() %>% 
    group_by(!!myenc) %>% 
    summarize(mean = mean(dogs_exp_prop),
              sd = sd(dogs_exp_prop),
              p25 = quantile(dogs_exp_prop, c(0.25)),
              p50 = quantile(dogs_exp_prop, c(0.50)),
              p75 = quantile(dogs_exp_prop, c(0.75))) %>% 
    ungroup()
}

seifa_cor <- function (seifa_index) {
  
  myenc <- enquo(seifa_index)
  
  SSC %>%
    st_drop_geometry() %>%
    select(!!myenc, dogs_exp_prop) %>%
    correlation(method = "kendall")
  
}

seifa_plot <- function (seifa_index) {
  
  model <- eval(substitute(lm(dogs_exp_prop ~ as.factor(seifa_index), 
                              data = SSC)))
  means <- estimate_means(model)
  
  myenc <- enquo(seifa_index)
  
  ggplot(SSC,
         aes(x = as.factor(!!myenc),
             y = dogs_exp_prop,
             fill = as.factor(!!myenc))) +
    geom_violin(alpha = 0.66) +
    geom_jitter2(width = 0.05, alpha = 0.5) +
    geom_line(data = means, aes(y = Mean, group = 1), size = 1) +
    geom_pointrange(data = means,
                    aes(y = Mean, ymin = CI_low, ymax = CI_high),
                    size = 1,
                    color = "white") + 
    scale_fill_brewer(palette = "BrBG") +
    theme_modern()
  
}
```

## IRSD {.tabset}

### Recalculated 

```{r}
seifa_means(IRSD_d)
seifa_cor(IRSD_d)
seifa_plot(IRSD_d)
```

### Original 

```{r}
seifa_means(IRSD_d_orig)
seifa_cor(IRSD_d_orig)
seifa_plot(IRSD_d_orig)
```

## IRSAD {.tabset}

### Recalculated 

```{r}
seifa_means(IRSAD_d)
seifa_cor(IRSAD_d)
seifa_plot(IRSAD_d)
```

### Original 

```{r}
seifa_means(IRSAD_d_orig)
seifa_cor(IRSAD_d_orig)
seifa_plot(IRSAD_d_orig)
```

## IER {.tabset}

### Recalculated 

```{r}
seifa_means(IER_d)
seifa_cor(IER_d)
seifa_plot(IER_d)
```

### Original 

```{r}
seifa_means(IER_d_orig)
seifa_cor(IER_d_orig)
seifa_plot(IER_d_orig)
```

## IEO {.tabset}

### Recalculated 

```{r}
seifa_means(IEO_d)
seifa_cor(IEO_d)
seifa_plot(IEO_d)
```

### Original 

```{r}
seifa_means(IEO_d_orig)
seifa_cor(IEO_d_orig)
seifa_plot(IEO_d_orig)
```

<!-- ----------------------------------------------------- -->

# Computing Environment

`r mu$session()`