---
title: "Dog-SEP"
subtitle: "Dog insurance"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    self_contained: true
    toc_float: yes
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

<!-- ------------------------------------------------------------ --> 

```{r r-setup, include=FALSE}
set.seed(12345)
options(scipen = 999)
options(max.print = "75")

library(pacman)
p_load(tidyverse, magrittr, scales, stringr,
       sjmisc, sjPlot, DT,
       sf, tmap,
       correlation, see, modelbased, parameters)

tmap_mode("view")
```

```{r knit-setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.width=8, fig.height=6, dpi=300, 
                      out.width="800px", out.height="600px")

knitr::opts_chunk$set(cache = FALSE,
                      prompt = FALSE,
                      tidy = FALSE,
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(width = 75)

mu <- Hmisc::markupSpecs$html
```

<!-- ------------------------------------------------------------ --> 

# Data 

Preparations in file `00.Rmd`

```{r}
dog_ownership_cost <- read_rds("data/dog_ownership_cost.Rds") %>% 
  select(-expensive) %>% 
  # mutate(expensive = ifelse(cost_compared_to_other_breeds == "Below average", 0, 1)) %>%
  mutate(expensive = ifelse(cost_compared_to_other_breeds == "Significantly above average", 1, 0)) %>%
  select(-cost_compared_to_other_breeds)

length(unique(dog_ownership_cost$SSC_NAME16))
length(unique(dog_ownership_cost$dog_breed))

SSC <- read_rds("data/geo/SSC.Rds")

length(unique(SSC$SSC_NAME16))

wide_insurance_n <- read_rds("data/wide_insurance_n.Rds")
wide_insurance_p <- read_rds("data/wide_insurance_p.Rds")
```

Important to remember, we have small amount of missings here!  

```{r echo=FALSE}
dog_ownership_cost %>% 
  frq(is.na(expensive))
```

Breeds:  

```{r echo=FALSE}
dog_ownership_cost %>% 
  filter(is.na(expensive)) %>% 
  frq(dog_breed, sort.frq = "desc")
```

## All Brisbane dogs combined 

```{r echo=FALSE, fig.width=12, fig.height=8}
my_colors <- RColorBrewer::brewer.pal(3, "BrBG")[c(1,3)]

dog_ownership_cost %>% 
  mutate(expensive = factor(expensive, levels = c(0, 1),
                            labels= c("No", "Yes"))) %>% 
  mutate(dog_breed = fct_lump(dog_breed, n = 50)) %>%
  group_by(dog_breed) %>% 
  mutate(total = n()) %>% 
  group_by(dog_breed, expensive) %>% 
  summarise(count = n(),
            total = first(total)) %>% 
  ggplot(aes(x = reorder(dog_breed, total) , 
             y = count,
             fill = expensive)) + 
  geom_col() +
  coord_flip() +
  ylab("Number of dogs") + xlab("") + 
  scale_fill_manual(values = my_colors) +
  theme_modern() +
  theme(panel.grid.major.x = element_line(colour = "grey80"),
        axis.text.y = element_text(size = 8))
```

## Dog counts

Summarizing all dogs, and expensive only.  

```{r}
dog_ownership_agg <- dog_ownership_cost %>% 
  group_by(SSC_NAME16) %>% 
  summarise(dogs_exp = sum(expensive, na.rm = TRUE))

SSC %<>% 
  left_join(dog_ownership_agg) %>%
  mutate(dogs_exp_prop = dogs_exp / dogs_total) %>% 
  relocate(geometry, .after = last_col())
```

## Proportion of expensive dogs

```{r echo=FALSE}
summary(SSC$dogs_exp_prop)
```

### Ranking

```{r echo=FALSE}
SSC %>% 
  st_drop_geometry() %>% 
  arrange(desc(dogs_exp_prop)) %>% 
  select(SSC_NAME16, URP, dogs_total, dogs_exp, dogs_exp_prop) %>% 
  mutate(dogs_exp_prop = number(dogs_exp_prop, accuracy = 0.01)) %>% 
  # slice(1:10) %>% 
  datatable()
```

### Map

```{r echo=FALSE}
tm_shape(SSC) +
  tm_bubbles(id = "SSC_NAME16", 
             size = "dogs_total", 
             col = "dogs_exp_prop",
             style = "quantile",
             palette = "BrBG",
             border.col = NA)
```

<!-- ------------------------------------------------------------ --> 

# Association with SEIFA

## Functions

```{r}
seifa_means <- function (seifa_index) {
  
  myenc <- enquo(seifa_index)
  
  SSC %>% 
    st_drop_geometry() %>% 
    group_by(!!myenc) %>% 
    summarize(mean = mean(dogs_exp_prop, na.rm = TRUE),
              sd = sd(dogs_exp_prop, na.rm = TRUE),
              p25 = quantile(dogs_exp_prop, c(0.25), na.rm = TRUE),
              p50 = quantile(dogs_exp_prop, c(0.50), na.rm = TRUE),
              p75 = quantile(dogs_exp_prop, c(0.75), na.rm = TRUE)) %>% 
    ungroup()
}

seifa_cor <- function (seifa_index) {
  
  myenc <- enquo(seifa_index)
  
  SSC %>%
    st_drop_geometry() %>%
    select(!!myenc, dogs_exp_prop) %>%
    mutate_if(is.factor, as.numeric) %>%
    correlation(method = "kendall")
  
}

seifa_plot <- function (seifa_index) {
  
  model <- eval(substitute(lm(dogs_exp_prop ~ seifa_index, 
                              data = SSC, na.action = na.omit)))
  means <- estimate_means(model)
  
  myenc <- enquo(seifa_index)
  
  ggplot(SSC,
         aes(x = !!myenc,
             y = dogs_exp_prop,
             fill = !!myenc)) +
    geom_violin(alpha = 0.66) +
    geom_jitter2(width = 0.05, alpha = 0.5) +
    geom_line(data = means, aes(y = Mean, group = 1), size = 1) +
    geom_pointrange(data = means,
                    aes(y = Mean, ymin = CI_low, ymax = CI_high),
                    size = 1,
                    color = "white") + 
    scale_fill_brewer(palette = "BrBG") +
    ylab("Proportion of expensive dogs") +
    theme_modern()
  
}
```

## IRSD {.tabset}

### Recalculated 

```{r message=FALSE, warning=FALSE}
seifa_means(IRSD_d)
seifa_cor(IRSD_d)
seifa_plot(IRSD_d)
```

### Original 

```{r message=FALSE, warning=FALSE}
seifa_means(IRSD_d_orig)
seifa_cor(IRSD_d_orig)
seifa_plot(IRSD_d_orig)
```

## IRSAD {.tabset}

### Recalculated 

```{r message=FALSE, warning=FALSE}
seifa_means(IRSAD_d)
seifa_cor(IRSAD_d)
seifa_plot(IRSAD_d)
```

### Original 

```{r message=FALSE, warning=FALSE}
seifa_means(IRSAD_d_orig)
seifa_cor(IRSAD_d_orig)
seifa_plot(IRSAD_d_orig)
```

## IER {.tabset}

### Recalculated 

```{r message=FALSE, warning=FALSE}
seifa_means(IER_d)
seifa_cor(IER_d)
seifa_plot(IER_d)
```

### Original 

```{r message=FALSE, warning=FALSE}
seifa_means(IER_d_orig)
seifa_cor(IER_d_orig)
seifa_plot(IER_d_orig)
```

## IEO {.tabset}

### Recalculated 

```{r message=FALSE, warning=FALSE}
seifa_means(IEO_d)
seifa_cor(IEO_d)
seifa_plot(IEO_d)
```

### Original 

```{r message=FALSE, warning=FALSE}
seifa_means(IEO_d_orig)
seifa_cor(IEO_d_orig)
seifa_plot(IEO_d_orig)
```

<!-- ----------------------------------------------------- -->

# PCA

<!-- https://stats.stackexchange.com/questions/53/pca-on-correlation-or-covariance -->

```{r}
data <- 
  # wide_insurance_n %>%
  wide_insurance_p %>%
  st_drop_geometry() %>% 
  select(alaskan_malamute:last_col()) %>% 
  as_tibble()

# View(cov(data))
```

```{r}
pca <- principal_components(data, 
                            standardize = FALSE,
                            n = 1)
pca

summary(pca)

plot(pca)

pca_results <- wide_insurance_p %>%
  st_drop_geometry() %>% 
  as_tibble() %>% 
  select(SSC_CODE16:caution) %>% 
  mutate(pca_raw = predict(pca)$Component_1, 
         pca = ntile(pca_raw, 10)) 
```

## `IRSD_d` 

```{r}
ggplot(pca_results, aes(x = IRSD, y = pca_raw)) + 
  geom_point()
```

```{r echo=FALSE}
pca_results %>% 
  frq(pca == IRSD_d)
```

```{r echo=FALSE}
plot_xtab(pca_results$IRSD_d, pca_results$pca, 
          margin = "row", bar.pos = "stack",
          show.summary = TRUE, coord.flip = TRUE)

tabz <- table(pca_results$IRSD_d, pca_results$pca)

assocplot(tabz,
          xlab = "IRSD_d", ylab = "pca")

mosaicplot(tabz,
           xlab = "IRSD_d", ylab = "pca")
```

## `IRSAD_d` 

```{r}
ggplot(pca_results, aes(x = IRSAD, y = pca_raw)) + 
  geom_point()
```

```{r echo=FALSE}
pca_results %>% 
  frq(pca == IRSAD_d)
```

```{r echo=FALSE}
plot_xtab(pca_results$IRSAD_d, pca_results$pca, 
          margin = "row", bar.pos = "stack",
          show.summary = TRUE, coord.flip = TRUE)

tabz <- table(pca_results$IRSAD_d, pca_results$IRSAD_d_orig)

assocplot(tabz,
          xlab = "IRSAD_d", ylab = "pca")

mosaicplot(tabz,
           xlab = "IRSAD_d", ylab = "pca")
```

## `IER_d` 

```{r}
ggplot(pca_results, aes(x = IER, y = pca_raw)) + 
  geom_point()
```

```{r echo=FALSE}
pca_results %>% 
  frq(pca == IER_d)
```

```{r echo=FALSE}
plot_xtab(pca_results$IER_d, pca_results$pca, 
          margin = "row", bar.pos = "stack",
          show.summary = TRUE, coord.flip = TRUE)

tabz <- table(pca_results$IER_d, pca_results$pca)

assocplot(tabz,
          xlab = "IER_d", ylab = "pca")

mosaicplot(tabz,
           xlab = "IER_d", ylab = "pca")
```

## `IEO_d` 

```{r}
ggplot(pca_results, aes(x = IEO, y = pca_raw)) + 
  geom_point()
```

```{r echo=FALSE}
pca_results %>% 
  frq(pca == IEO_d)
```

```{r echo=FALSE}
plot_xtab(pca_results$IEO_d, pca_results$pca, 
          margin = "row", bar.pos = "stack",
          show.summary = TRUE, coord.flip = TRUE)

tabz <- table(pca_results$IEO_d, pca_results$pca)

assocplot(tabz,
          xlab = "IEO_d", ylab = "IEO_d_orig")

mosaicplot(tabz,
           xlab = "IEO_d", ylab = "IEO_d_orig")
```

<!-- ----------------------------------------------------- -->

# Clustering I

```{r}
data <- 
  wide_insurance_p %>%
  st_drop_geometry() %>% 
  select(alaskan_malamute:last_col()) %>% 
  as_tibble()
```

## No of clusters

```{r}
n <- n_clusters(data, package = c("easystats", "NbClust", "mclust"))
n
plot(n)
```

## K-Means

```{r}
rez_kmeans <- cluster_analysis(data, n = 2, method = "kmeans")

rez_kmeans

plot(rez_kmeans)

plot(summary(rez_kmeans))
```

```{r}
cluster_results <- wide_insurance_p %>%
  st_drop_geometry() %>% 
  as_tibble() %>% 
  mutate(cluster = predict(rez_kmeans)) 

frq(cluster_results, cluster)

aggregate(data = cluster_results, french_bulldog ~ cluster, mean)
aggregate(data = cluster_results, rottweiler ~ cluster, mean)
```

```{r echo=FALSE}
cluster_results_map <- 
  left_join(SSC, 
            cluster_results %>% 
              select(SSC_CODE16, cluster) %>% 
              mutate(cluster = factor(cluster)))

tm_shape(cluster_results_map) +
  tm_bubbles(id = "SSC_NAME16", 
             size = "dogs_total", 
             col = "cluster",
             # palette = "BrBG",
             border.col = NA)
```

## Hierarchical Clustering

```{r}
rez_hclust <- cluster_analysis(data, n = 2, method = "hclust")

rez_hclust

plot(rez_hclust)
```

## Hierarchical K-Means

```{r}
rez_hkmeans <- cluster_analysis(data, n = 2, method = "hkmeans")

rez_hkmeans

plot(rez_hkmeans)
```

## K-Medoids (PAM)

```{r}
rez_pam <- cluster_analysis(data, n = 2, method = "pam")

rez_pam

plot(rez_pam)
```

## Bootstrapped Hierarchical Clustering

```{r eval=FALSE, include=FALSE}
rez_hclust2 <- cluster_analysis(data, 
                                n = NULL, 
                                method = "hclust", 
                                iterations = 500,
                                ci = 0.90)

rez_hclust2

plot(rez_hclust2)
```

## DBSCAN

```{r}
eps <- n_clusters_dbscan(data, min_size = 0.01) 

eps

plot(eps)

rez_dbscan <- cluster_analysis(data, method = "dbscan", dbscan_eps = 4.5)

rez_dbscan

plot(rez_dbscan)
```

## Hierarchical K-Means

```{r}
rez_hdbscan <- cluster_analysis(data, method = "hdbscan")

rez_hdbscan

# plot(rez_hdbscan)
```

## K-Medoids with estimation of number of clusters (pamk)

```{r}
rez_pamk <- cluster_analysis(data, method = "pamk")

rez_pamk 

plot(rez_pamk)
```

## Mixture

```{r}
p_load(mclust)

rez_mixture <- cluster_analysis(data, method = "mixture")

rez_mixture

plot(rez_mixture)
```

## Metaclustering

```{r}
list_of_results <- list(rez_kmeans, rez_hclust, rez_hkmeans, rez_pam,
                        # rez_hclust2, 
                        rez_dbscan, rez_hdbscan, rez_mixture)

probability_matrix <- cluster_meta(list_of_results)

heatmap(probability_matrix, scale = "none", 
        col = grDevices::hcl.colors(256, palette = "inferno"))
```

<!-- ----------------------------------------------------- -->

# Clustering II

```{r}
data <- 
  wide_insurance_p %>%
  st_drop_geometry() %>% 
  select(alaskan_malamute:last_col()) %>% 
  as_tibble()
```

```{r}
p_load(mclust)

BIC <- mclustBIC(data, G = seq(from = 2, to = 9))

plot(BIC)

# BIC
summary(BIC)

mod1 <- Mclust(data, x = BIC)
mod1
summary(mod1, parameters = TRUE)

frq(mod1$classification)

plot(mod1, what = "classification")

# ICL <- mclustICL(data, G = seq(from = 2, to = 9))
# summary(ICL)
# plot(ICL)

# LRT <- mclustBootstrapLRT(data, modelName = "VEI")
# summary(LRT)
# plot(LRT)
```


```{r}
cluster_results <- wide_insurance_p %>%
  st_drop_geometry() %>% 
  as_tibble() %>% 
  mutate(cluster = factor(mod1$classification)) 

aggregate(data = cluster_results, french_bulldog ~ cluster, mean)
aggregate(data = cluster_results, rottweiler ~ cluster, mean)
```

```{r echo=FALSE}
cluster_results_map <- 
  left_join(SSC, 
            cluster_results %>% 
              select(SSC_CODE16, cluster) %>% 
              mutate(cluster = factor(cluster)))

tm_shape(cluster_results_map) +
  tm_bubbles(id = "SSC_NAME16", 
             size = "dogs_total", 
             col = "cluster",
             # palette = "BrBG",
             border.col = NA)
```

<!-- ----------------------------------------------------- -->

# Computing Environment

`r mu$session()`