---
title: "Dog-SEP"
subtitle: "Machine learning: support vector machine"
pagetitle: "Dog-SEP: SVM"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    self_contained: true
    toc_float: yes
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

<!-- ------------------------------------------------------------ --> 

```{r r-setup, include=FALSE}
set.seed(12345)
options(scipen = 999)
options(max.print = "75")

library(pacman)
p_load(tidyverse, magrittr, DT,
       caret, e1071, kernlab)

import::from("sjmisc", "frq")
```

```{r knit-setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.width = 8, fig.height = 6, dpi = 300, 
                      out.width = "800px", out.height = "600px")

knitr::opts_chunk$set(cache = FALSE,
                      prompt = FALSE,
                      tidy = FALSE,
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(width = 75)

mu <- Hmisc::markupSpecs$html
```

```{r include=FALSE}
# function to calculate the mean absolute error
MAE <- function(actual, predicted) {
  mean(abs(actual - predicted))  
}
```

<!-- ------------------------------------------------------------ --> 

# Background 

Using info from chapter 7 of *Machine Learning with R* [book](https://www.packtpub.com/product/machine-learning-with-r-third-edition/9781788295864), `caret` package [manual](https://topepo.github.io/caret/index.html), and `e1071` package [vignette](https://cran.r-project.org/web/packages/e1071/).  

<!-- ------------------------------------------------------------ --> 

# Analysis with `e1071`

## Data 

```{r include=FALSE}
training <- read_rds("data/training.Rds")
testing <- read_rds("data/testing.Rds")
```

## Training 

```{r}
set.seed(12345)
m_svm <- svm(outcome ~ ., data = training,
             type = "eps-regression",
             kernel = "radial")
```

```{r echo=FALSE}
summary(m_svm)
```

## Prediction

```{r}
p_svm <- predict(m_svm, testing[-1])
```

Correlation:  

```{r}
cor(p_svm, testing$outcome)
```

MAE between predicted and actual values:  

```{r}
MAE(p_svm, testing$outcome)
```

Distro of predicted (red) nd actual (green): 

```{r echo=FALSE}
# summary(p_svm)
# summary(testing$outcome)

ggplot() +
  geom_density(data = as.data.frame(p_svm), aes(x = p_svm), col = "firebrick") +
  geom_density(data = testing, aes(x = outcome), col = "forestgreen") 
```

<!-- ------------------------------------------------------------ --> 

# Analysis with `kernlab`

## Tuning

```{r}
modelLookup("svmRadial")
```

Using grid of parameters to explore:  

```{r}
grid_tu <-  expand.grid(sigma = seq(0.00001, 0.0002, 0.00001),
                        C = seq(10, 20, 1))

nrow(grid_tu)
```

```{r}
p_load(doParallel)

cl <- makePSOCKcluster(parallel::detectCores())
registerDoParallel(cl)

set.seed(12345)
m_svm_tu <- train(outcome ~ .,
                    method = "svmRadial",
                    data = training,
                    # weights = training_weights, 
                    metric = "RMSE",
                    # metric = "Rsquared",
                    tuneGrid = grid_tu)

stopCluster(cl)
p_unload(doParallel)
```

```{r echo=FALSE}
# summary(m_svm_tu)
ggplot(m_svm_tu)
```

```{r}
m_svm_tu$finalModel
```

```{r}
p_svm_tu <- predict(m_svm_tu, testing[-1])
```

Correlation:  

```{r}
cor(p_svm_tu, testing$outcome)
```

MAE between predicted and actual values:  

```{r}
MAE(p_svm_tu, testing$outcome)
```

Distro of predicted (red) nd actual (green): 

```{r echo=FALSE}
# summary(p_svm)
# summary(testing$outcome)

ggplot() +
  geom_density(data = as.data.frame(p_svm_tu), aes(x = p_svm_tu), col = "firebrick") +
  geom_density(data = testing, aes(x = outcome), col = "forestgreen") 
```

