---
title: "Principal Component Analysis"
author: "Dataset data"
output:
  html_document: default
---
This dataset contains 180 individuals and 19 variables.

- - -

```{r, echo = FALSE}
library(FactoMineR)
load('C:/external/FUN_BCC-animals/Workspace.RData')
```
### 1. Study of the outliers
The analysis of the graphs does not detect any outlier.

- - -

### 2. Inertia distribution
The inertia of the first dimensions shows if there are strong relationships between variables and suggests the number of dimensions that should be studied.

The first two dimensions of analyse express **96.08%** of the total dataset inertia ; that means that 96.08% of the individuals (or variables) cloud total variability is explained by the plane.
This percentage is particularly high and thus the first plane perfectly represents the data variability.
This value is strongly greater than the reference value that equals **17.74%**, the variability explained by this plane is thus highly significant
(the reference value is the 0.95-quantile of the inertia percentages distribution obtained by simulating 2904 data tables of equivalent size on the basis of a normal distribution).

From these observations, it is absolutely not necessary to interpret the next dimensions.

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
par(mar = c(2.6, 4.1, 1.1, 2.1))
ggplot2::ggplot(cbind.data.frame(x=1:nrow(res$eig),y=res$eig[,2])) + ggplot2::aes(x=x, y=y)+ ggplot2::geom_col(fill="blue") + ggplot2::xlab("Dimension") + ggplot2::ylab("Percentage of variance") + ggplot2::ggtitle("Decomposition of the total inertia") + ggplot2::theme_light() + ggplot2::theme(plot.title = ggplot2::element_text(hjust =0.5)) + ggplot2::scale_x_continuous(breaks=1:nrow(res$eig))
```

**Figure 2 - Decomposition of the total inertia**
*The first factor is largely dominant: it expresses itself 90.18% of the data variability.*
*Note that in such a case, the variability related to the other components might be meaningless, despite of a high percentage.*

An estimation of the right number of axis to interpret suggests to restrict the analysis to the description of the first 1 axis.
These axis present an amount of inertia greater than those obtained by the 0.95-quantile of random distributions (90.18% against 9.41%).
This observation suggests that only this axis is carrying a real information.
As a consequence, the description will stand to these axis.

- - -

### 3. Description of the dimension 1

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("59", "20", "34", "83", "77", "165", "60")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = c(1,2), choix = 'ind', invisible = 'quali', title = '', cex = cex)
```

**Figure 3.1 - Individuals factor map (PCA)**
*The labeled individuals are those with the higher contribution to the plane construction.*

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("maltese", "german_shepherd", "rottweiler", "french_bulldog", 
"british_bulldog", "irish_wolfhound", "dogue_de_bordeaux")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = 1:2, choix = 'var', title = '', cex = cex)
```

**Figure 3.2 - Variables factor map (PCA)**
*The labeled variables are those the best shown on the plane.*

* * *

The **dimension 1** opposes individuals such as *59*, *20*, *165*, *83* and *77* (to the right of the graph, characterized by a strongly positive coordinate on the axis)
to individuals such as ** and *34* (to the left of the graph, characterized by a strongly negative coordinate on the axis).

The group in which the individuals *165*, *83* and *77* stand (characterized by a positive coordinate on the axis) is sharing :

- high values for the variables *maltese*, *german_shepherd*, *rottweiler*, *british_bulldog*, *french_bulldog*, *akita*, *irish_wolfhound*, *yorkshire_terrier*, *samoyed* and *chinese_crested_dog* (variables are sorted from the strongest).

The group in which the individuals *59* and *20* stand (characterized by a positive coordinate on the axis) is sharing :

- high values for the variables *german_shepherd*, *maltese*, *rottweiler*, *dogue_de_bordeaux*, *french_bulldog*, *chinese_crested_dog*, *irish_wolfhound* and *british_bulldog* (variables are sorted from the strongest).

The group in which the individual *34* stands (characterized by a negative coordinate on the axis) is sharing :

- low values for variables like *maltese*, *german_shepherd*, *rottweiler*, *british_bulldog*, *french_bulldog*, *irish_wolfhound*, *akita*, *yorkshire_terrier*, *dogue_de_bordeaux* and *chinese_crested_dog* (variables are sorted from the weakest).

Note that the variable *maltese* is highly correlated with this dimension (correlation of 0.99). This variable could therefore summarize itself the dimension 1.

- - -

### 4. Classification

```{r, echo = FALSE}
res.hcpc = HCPC(res, nb.clust = -1, graph = FALSE)
```

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("59", "20", "34", "83", "77", "165", "60")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.HCPC(res.hcpc, choice = 'map', draw.tree = FALSE, select = drawn, title = '')
```

**Figure 4 - Ascending Hierarchical Classification of the individuals.**
*The classification made on individuals reveals 3 clusters.*


The **cluster 1** is made of individuals such as *34* and *60*. This group is characterized by :

- low values for variables like *maltese*, *german_shepherd*, *rottweiler*, *british_bulldog*, *irish_wolfhound*, *dogue_de_bordeaux*, *akita*, *yorkshire_terrier*, *chow_chow* and *samoyed* (variables are sorted from the weakest).

The **cluster 2** is made of individuals such as *77* and *83*. This group is characterized by :

- high values for the variables *maltese*, *german_shepherd*, *rottweiler*, *british_bulldog*, *irish_wolfhound*, *akita*, *chow_chow* and *dogue_de_bordeaux* (variables are sorted from the strongest).

The **cluster 3** is made of individuals such as *20*, *59* and *165*. This group is characterized by :

- high values for the variables *maltese*, *german_shepherd*, *rottweiler*, *dogue_de_bordeaux*, *yorkshire_terrier*, *british_bulldog*, *samoyed*, *french_bulldog* and *canadian_eskimo_dog* (variables are sorted from the strongest).

- - -

## Annexes



